---
title: "Kriging"
author: "Nothando Maloleka"
date: "2023-10-08"
output:
  word_document: default
  html_document: default
---
```{r}

#The distribution of Rainfall in Switzerland
library(geoR)
points(sic.100, borders=sic.borders,col="green")
points(sic.367, borders=sic.borders,col="red",add=TRUE)
library(geoR)
plot.geodata(sic.100,bor=sic.borders)

#Descriptive Statistics 
library(geoR)
library(fields)
vario.b<- variog(sic.100,option =c ("bin", "cloud", "smooth"),
bin.cloud=TRUE)

```
`````{r}
library("geoR")
vario.ex<- variog(sic.100, bin.cloud=TRUE)
plot(vario.ex,main="")
lines.variomodel(cov.model="spher",cov.pars=c(15000,50),
nug=0,max.dist=300)
````


```{r}
library(geoR)
pred.grid <- expand.grid(seq(0,350, l=51),seq (0,220, l=51))
rgb.palette <- colorRampPalette(c("blue", "lightblue",
"orange", "red"),space = "rgb")
kc<- krige.conv(sic.100, loc = pred.grid,
krige=krige.control(cov.model="spherical",cov.pars=c(15000,50)))
image(kc, loc = pred.grid,col =rgb.palette(20) ,xlab="Coord X",
ylab="Coord Y",borders=sic.borders,main="Estimation")
image(kc, krige.var,loc = pred.grid,col=rgb.palette(20),
xlab="Coord X",ylab="Coord Y",borders=sic.borders,
main="Kriging variance")

```
````{r}
#Estimated and Observed values
library("geoR")
kc1<- krige.conv(sic.100, loc = sic.100$coords,
krige=krige.control(cov.model="spherical",cov.pars=c(16000,47)))
kc2<- krige.conv(sic.100, loc = sic.367$coords,
krige=krige.control(cov.model="spherical",cov.pars=c(16000,47)))
plot(sic.100$data,kc1$predict,xlab="Observed",ylab="Estimated",
main="Control sample")
abline(a=0,b=1,col="red")
plot(sic.367$data,kc2$predict,,xlab="Observed",ylab="Estimated",
main="Control")
abline(a=0,b=1,col="red")    

`````

````{r}
#Descriptive statistics after Box-Cox Transformation
library("geoR")
plot.geodata(sic.100,bor=sic.borders,lambda=0.5)
````
``````{r}
library("geoR")
vario.ext <- variog(sic.100, option = "bin", lambda = 0.5)
plot(vario.ext)
lines.variomodel(cov.m = "mat", cov.p = c(105, 36), nug = 6.9,
                 max.dist = 300, kappa = 1, lty = 1)
``````



`````{r}
library(geoR)
kct<- krige.conv(sic.100, loc = pred.grid,
krige=krige.control(cov.model="matern",cov.pars=c(105, 36),
kappa=1,nugget=6.9,lambda=0.5))
pred.grid <- expand.grid(seq(0,350, l=51),seq (0,220, l=51))
rgb.palette <- colorRampPalette(c("blue", "lightblue",
"orange", "red"),space = "rgb")
image(kct, loc = pred.grid,col =rgb.palette(20) , xlab="Coord X",
ylab="Coord Y",borders=sic.borders,main="Estimation") 
krige.var1<- kct$krige.var
image(kct, krige.var1,loc = pred.grid,col =rgb.palette(20) ,
xlab="Coord X",ylab="Coord Y",borders=sic.borders,
main="Kriging variance")

````
`


```````{r}
#Anistrophic Kriging
library("geoR")
kc_aniso <- krige.conv(sic.100, loc = pred.grid,
                       krige = krige.control(cov.model = "spherical", cov.pars = c(15000, 50),
                                             aniso = c(1, 2)))

# Visualization
image(kc_aniso, loc = pred.grid, col = rgb.palette(20),
      xlab = "Coord X", ylab = "Coord Y", borders = sic.borders,
      main = "Anisotropic Kriging")

#Universal Kriging

library(geoR)
data(SIC)
points(sic.100, borders = sic.borders, pch=20, cex.max=3)
max.dist <- 220
sic.bin<-variog(sic.100,option="bin",estimator.type="classical",
max.dist=max.dist)

```````
```````{r}

krige.control(cov.model = "gaussian", cov.pars = c(15000, 50))
krige.control(cov.model = "spherical", cov.pars = c(20000, 30))
krige.control(cov.model = "spherical", cov.pars = c(15000, 50), aniso = c(1, 2))
library("geoR")
kc_aniso <- krige.conv(sic.100, loc = pred.grid,
                       krige = krige.control(cov.model = "spherical", cov.pars = c(15000, 50),
                                             aniso = c(1, 2)))

#Kriging with Gaussian Covariance Model

library("geoR")
kc_gaussian <- krige.conv(sic.100, loc = pred.grid,
                          krige = krige.control(cov.model = "gaussian", cov.pars = c(15000, 50)))

# Visualization
image(kc_gaussian, loc = pred.grid, col = rgb.palette(20),
      xlab = "Coord X", ylab = "Coord Y", borders = sic.borders,
      main = "Kriging with Gaussian Covariance Model")

`````
`````{r}
library(geoR)
kct1<- krige.conv(sic.100, loc = sic.100$coords,
krige=krige.control(cov.model="spherical",cov.pars=c(16000,47),
kappa=1,nugget=6.9,lambda=0.5))
kct2<- krige.conv(sic.100, loc = sic.367$coords,
krige=krige.control(cov.model="spherical",cov.pars=c(16000,47),
kappa=1,nugget=6.9,lambda=0.5))
plot(sic.100$data,kct1$predict,xlab="Observed",ylab="Estimated",
main="Sample")
abline(a=0,b=1,col="red")
plot(sic.367$data,kct2$predict,,xlab="Observed",ylab="Estimated",
main="Control sample")
abline(a=0.0009,b=8.2,col="red")



````
`````{r}

# Assuming you have a validation dataset in 'validation_data' with columns 'observed' and 'location'
# Also assuming 'kct1' and 'kct2' are the kriging results from the previous code

# Extract observed values from validation dataset for sample and control
observed_sample <- sic.100$data
observed_control <- sic.367$data  

# Extract kriging predictions for sample and control
predicted_sample <- kct1$predict
predicted_control <- kct2$predict

# Calculate RMSE for sample and control
rmse_sample <- sqrt(mean((observed_sample - predicted_sample)^2))
rmse_control <- sqrt(mean((observed_control - predicted_control)^2))

# Calculate covariance for sample and control
#covariance_sample <- cov(observed_sample, predicted_sample)[1, 2]
#covariance_control <- cov(observed_control, predicted_control)[1, 2]

# Print RMSE and covariance
cat("RMSE for Sample Data:", rmse_sample, "\n")
#cat("Covariance for Sample Data:", covariance_sample, "\n")

cat("RMSE for Control Sample Data:", rmse_control, "\n")
#cat("Covariance for Control Sample Data:", covariance_control, "\n")

````
````{r}
# Assuming you have a validation dataset in 'validation_data' with columns 'observed' and 'location'
# Also assuming 'kct1' and 'kct2' are the kriging results from the previous code

# Extract observed values from validation dataset for sample and control
observed_sample <- sic.100$data
observed_control <- sic.367$data  

# Extract kriging predictions for sample and control
predicted_sample <- kct1$predict
predicted_control <- kct2$predict

# Calculate RMSE for sample and control
rmse_sample <- sqrt(mean((observed_sample - predicted_sample)^2))
rmse_control <- sqrt(mean((observed_control - predicted_control)^2))

# Calculate covariance for sample and control
#covariance_sample <- cov(observed_sample, predicted_sample)[1, 2]
#covariance_control <- cov(observed_control, predicted_control)[1, 2]

# Print RMSE and covariance
cat("RMSE for Sample Data:", rmse_sample, "\n")
#cat("Covariance for Sample Data:", covariance_sample, "\n")

cat("RMSE for Control Sample Data:", rmse_control, "\n")
#cat("Covariance for Control Sample Data:", covariance_control, "\n")
```
